Increasing the mortality gap? Social inequality in mortality among elderly and adults in Northern Sweden 1851--2013\footnote{Session "Health, morbidity, and mortality" at the Nordic Demographic Symposium 14--16 June 2017, Turku, Finland}

### 2017-10-12 cape.Rnw
add slide with deaths by period and sex

\end{document}

>>>>

\begin{frame}{Crude statistics}

<<crude>>=
dat <- with(vb0Tab, aggregate(list(event = event, exposure = exposure), 
                              by = list(period = period, sex = sex), 
                              FUN = sum))
males <- dat[dat$sex == "male", c("period", "event", "exposure")]
females <- dat[dat$sex == "female", c("period", "event", "exposure")]
oldpar <- par(mfrow = c(1, 2))
x <- c(1812.5, 1837.5, 1862.5, 1888.5, 1912.5, 1937.5, 1987.5, 1982.5, 1987.5, 
       1992.5, 1997.5, 2004.5, 2010.5)
plot(x, males$event, type = "b", col = "blue", main = "Number of deaths by period",
     ylim = c(0, 7500), ylab = "No. of deaths", xlab = "Time period", lty = 1, axes = FALSE)
axis(1, at = c(1801, 1826, 1851, 1876, 1901, 1926, 1951, 1976, 2014), las = 2)
axis(2, las = 1)
box()
lines(x, females$event, type = "b", col = "red", lty = 2)
legend("bottomright", legend = c("Males", "Females"), lty = c(1, 2), col = c("blue", "red"))
plot(x, males$event / males$exposure * 1000, type = "b", col = "blue", 
     main = "Crude death rates, ages 40-89", ylim = c(0, 30), xlim = c(1800, 2014),
     xlab = "Time period", ylab = "Deaths per 1000", axes = FALSE)
axis(1, at = c(1801, 1826, 1851, 1876, 1901, 1926, 1951, 1976, 2014), las = 2)
axis(2, las = 1)
box()
lines(x, females$event / females$exposure * 1000, type = "b", col = "red")
legend("bottomright", legend = c("Males", "Females"), lty = c(1, 2), col = c("blue", "red"))
@

\end{frame}
\end{document}
